### 2015-02-26 paper/intro.tex
removed REDUX
title, abstract, first round of corrections

\section*{Overview}
Rust makes strong claims about ensuring memory safety without garbage collection.
We would like to prove that those claims are true.
To that end, we use a small model of Rust, called Patina, that characterizes the key
features of Rust under the most suspicion, namely its memory management.

Rust memory management has two goals:
no memory should ever be orphaned and no uninitialized memory should ever be read.
Rust tries to achieve this by maintaining two invariants:
all memory has a unique owner responsible for deallocating it and
no memory is ever simultaneously aliased and mutable.
These invariants simplify the situation enough so that
Rust only needs to track which L-values are uninitialized and which have been borrowed.
Ownership tracking of heap memory is managed by unique pointers.
Safe aliasing is managed by borrowed references.

The Patina model has three layers.
The innermost layer deals with L-values and ensures no uninitialized pointer
ever needs to be dereferenced. It also ensures that initialization data correctly
models the runtime memory.
The middle layer deals with R-values.
It ensures that an L-values used do not violate initialization or borrow restrictions.
It also ensures that borrowed references are not created unless they are safe.
The outer layer deals with statements.
It mostly just chains together the guarantees of the L-value and R-value layers.
However, it is also responsible for ensuring no orphans would be created when deallocating memory.

@@

Unique pointers can also be freed if they would be orphaned by assignment.
\begin{verbatim}
{
  let x: ~~int = ~~3;
  *x = ~1; // The ~3 that *x points too would be orphaned here. It is freed instead.
}
\end{verbatim}

@@

\subsection*{Unique Pointers in Patina}
To check that the Rust compiler does this correctly, Patina uses a shallow free statement.
The only way to deallocate a unique pointer in Patina is with these explicit frees.
\begin{verbatim}
{
  let x: ~~int = ~~3;
  //free(x); // ERROR! would orphan *x
  // The proper way
  free(*x);
  free(x);
}
\end{verbatim}
However, the free statement is only valid for initialized pointers, which prevents double frees.
\begin{verbatim}
{
  let x: ~int = ~3;
  free(x); // *x is now deallocated. x is uninitialized
  free(x); // ERROR! x is not initialized
}
\end{verbatim}
Finally, the free statement requires that the pointer is unaliased, preventing dangling pointers.
\begin{verbatim}
{
  let x: ~int = ~3;
  let y: &int = &*x; // Create an immutable borrowed reference to *x
  free(x); // ERROR! would make y a dangling pointer
}
\end{verbatim}

\subsection*{Borrowed Referenced REDUX}
Often, particularly in functions, we want to use some memory, but do not intend to free it.
That is, we want to borrow access to memory owned by something else.
This is fine for copyable values, like integers, 
but is tedious for noncopyable values such as unique pointers.
To avoid freeing a unique pointer argument at the end of the function body,
it must be returned from the function.
Effectively, the programmer must thread noncopyable values through their function calls.
\begin{verbatim}
fn foo(x: ~int) -> (int, ~int) {
  return (*x + 1, x)
}

@@

{
  let x: ~int = ~1;
  let y: int = foo(&*x);
  // x = ~1, y = 2
}
\end{verbatim}

@@

%\subsection*{Borrowed References}
%Borrowed references are a means to provide safe, temporary aliases in Rust.
%They do so by enforcing two invariants.
%The first is that references are always valid.
%\begin{verbatim}
%{
%  let x: int; // An uninitialized stack integer
%  let y: &int = &x; // ERROR! *y would point to uninitialized memory
%}
%\end{verbatim}
%This also extends to cases where the referent falls out of scope too soon.
%\begin{verbatim}
%{
%  let y: &int;
%  {
%    let x: int = 3;
%    y = &x; // ERROR! x does not live long enough
%  } // x would be freed here
%  // *y would point to unallocated memory here
%}
%\end{verbatim}
%The second invariant is to prevent any memory from being simultaneously aliased and mutable.
%This is to avoid situtations like the following:
%\begin{verbatim}
%{
%  let x: Option<int> = Some(3);
%  let y: &int = match x {
%    Some(ref z) => z,
%    None => fail!(), // Impossible
%  }
%  // If x is mutable, then the payload 3 is both aliased and mutable
%  x = None // The payload is now uninitialized so *y is a dangling pointer
%}
%\end{verbatim}
%Therefore immutable borrows freeze the underlying data.
%It cannot be mutated by anything while the borrow exists.
%However, it is freely copyable.
%\begin{verbatim}
%{
%  let x: int = 3;
%  let y: &int = &x;
%  let z: &int = y; // Both y and z are readable
%  x = 2; // ERROR! x is frozen by y
%}
%\end{verbatim}
%Conversely, mutable borrows forbid all other access to the underlying data.
%It does allow the data to be mutated, but it is an affine datatype like unique pointers.
%\begin{verbatim}
%{
%  let x: int = 3;
%  let y: &mut int = &mut x;
%  //let z: int = x; // ERROR! cannot use x while mutable borrowed
%  let z: int = *y; // OK
%
%  let w: &mut int = y; // Moves y into w. y is not uninitialized
%  z = *y; // ERROR! y is uninitialized
%}
%\end{verbatim}
%As with unique pointers, these checks are all static.

@@

$\own{\gt}$ is a unique pointer to a $\gt$.
$\tyref{\lt}{q}{\gt}$ is a borrowed reference to a $\gt$
providing mutability guarantee $q$ for lifetime $\lt$.
$\subvar{\gt}{i}{n}$ is a variant type.
$\subrec{\gt}{i}{n}$ is a tuple type.
Finally, $\tyfix{X}{\gt}$ is a recursive type.

>>>>

\section*{Abstract}
Rust is a new systems language that uses some advanced type system features,
specifically affine types and regions, to statically guarantee memory
safety and eliminate the need for a garbage collector.
While each individual addition to the type system is well understood
in isolation and are known to be sound, the combined system is not.
Furthermore, Rust uses a novel checking scheme for its regions, known
as the Borrow Checker, that is not known to be correct.
Since Rust's goal is to be a safer alternative to C/C++, we should
ensure that this safety scheme actually works. We present a formal
semantics that camptures the key features relevant to memory safety,
unique pointers and borrowed references, specifies how they guarantee
memory safety, and describes the operation of the Borrow Checker.
We used this model to prove the soudness of some core operations
and justify the conjecture that the model, as a whole, is sound.
Additionally, our model provides a syntactic version of the Borrow
Checker, which may be more understandable than the non-syntactic version
in Rust.

\section*{Overview}
The Rust Programming Language~\cite{rust} makes strong claims about 
ensuring memory safety without garbage collection.
We would like to prove that those claims are true.
To that end, we use a small model of Rust, called Patina, that characterizes the key
features of Rust under the most suspicion, namely its memory management.

Rust memory management has two goals:
no memory should ever become unreachable without being deallocated
and no uninitialized memory should ever be read.
Rust tries to achieve this by maintaining two invariants:
all memory has a unique owner responsible for deallocating it and
no memory is ever simultaneously aliased and mutable.
These invariants simplify the situation enough so that
Rust only needs to track which L-values are uninitialized and which have been borrowed.
Ownership tracking of heap memory is managed by unique pointers.
Safe aliasing is managed by borrowed references.

The Patina model has three layers.
The innermost layer deals with L-values and ensures 
no uninitialized data is ever used, which includes ensuring
that uninitialized pointers and null pointers are never dereferenced.
It also ensures that compile-time initialization information correctly
models the runtime memory.
The middle layer deals with R-values.
It ensures that any L-values used do not violate initialization or borrow restrictions.
It also ensures that borrowed references are not created unless they are safe.
The outer layer deals with statements.
It mostly just chains together the guarantees of the L-value and R-value layers.
However, it is also responsible for ensuring no leaks would be created when deallocating memory.

@@

Unique pointers can also be freed if they would be leaked by assignment.
\begin{verbatim}
{
  let x: ~~int = ~~3;
  *x = ~1; // The ~3 that *x points too would be leaked here. It is freed instead.
}
\end{verbatim}

@@

\subsection*{Unique Pointers in Patina}
When unique pointers fall out of scope in Rust, the compiler inserts code at the end
of the block that will actually free the allocated heap memory. This code will
traverse the entire owned data structure and free memory without leaking anything.
In the case of recursive data structures, this code will be a recursive function.
Rather than model such complex behavior as a primitive in Patina, we will use only
a shallow free statement. Aside from making the model simpler, this also provides us
with a means of checking that the code Rust inserts to free memory is correct,
i.e.\ by encoding it in terms of Patina's shallow frees.
The only way to deallocate a unique pointer in Patina is with these explicit frees.
\begin{verbatim}
{
  let x: ~~int = ~~3;
  //free(x); // ERROR! would leak *x
  // The proper way
  free(*x);
  free(x);
}
\end{verbatim}
However, the free statement is only valid for initialized pointers, which prevents double frees.
\begin{verbatim}
{
  let x: ~int = ~3;
  free(x); // *x is now deallocated. x is uninitialized
  free(x); // ERROR! x is not initialized
}
\end{verbatim}
Finally, the free statement requires that the pointer is unaliased, preventing dangling pointers.
\begin{verbatim}
{
  let x: ~int = ~3;
  // Create an immutable borrowed reference to *x with &
  // Immutable borrowed references to some type A are denoted by &A
  let y: &int = &*x;
  free(x); // ERROR! would make y a dangling pointer
}
\end{verbatim}

\subsection*{Borrowed References}
Often, particularly in functions, we want to use some memory, but do not intend to free it.
That is, we want to borrow access to memory owned by something else.
This is fine for copyable values, like integers, 
but is tedious for noncopyable values such as unique pointers.
To avoid freeing a unique pointer argument at the end of the function body,
it must be returned from the function.
Effectively, the programmer must thread noncopyable values through their function calls.
\begin{verbatim}
fn foo(x: ~int) -> (int, ~int) {
  return (*x + 1, x)
}

@@

{
  let x: ~int = ~1;
  let y: int = foo(&*x); // Rust lets us write foo(x) here and will insert the &* for us
  // x = ~1, y = 2
}
\end{verbatim}

@@

@@

$\own{\gt}$ is a unique pointer to a $\gt$.
$\tyref{\lt}{q}{\gt}$ is a borrowed reference to a $\gt$
providing mutability guarantee $q$ for lifetime $\lt$, which is Rust's name
for regions associated with stack frames.
$\subvar{\gt}{i}{n}$ is a discriminated union or a variant type.
$\subrec{\gt}{i}{n}$ is a tuple type.
Finally, $\tyfix{X}{\gt}$ is a recursive type.

### 2015-02-26 paper/paths.tex
removed old commented out text
added the reachability constraint to the well-formed judgment for heaps
added unique pointer uniqueness constraint to heap typing
title, abstract, first round of corrections

%Unsurprisingly, path typing is unique.
%\begin{lem}[Path Type Uniqueness]
%  If $\tc{\gG}{\lv{x}{p}}{\gt}$ and $\tc{\gG}{\lv{x}{p}}{\gt'}$ then $\gt = \gt' $.
%\end{lem}
%
%\begin{proof}[Path Type Uniqueness]
%  We will use induction on the derivation of $\tc{\gG}{\lv{x}{p}}{\gt}$.
%  \begin{itemize}
%    \item[\textsc{PT-BASE}] Then $p = \base$. \\
%      By inversion, $\gG(x)=\gt$. \\
%      There is only one possible rule (\textsc{PT-BASE}) for 
%      deriving $\tc{\gG}{\lv{x}{\base}}{\gt'}$ \\
%      Ergo, by inversion, $\gG(x)=\gt'$. \\
%      Thus, $\gt = \gt'$.
%    \item[\textsc{PT-DEOWN}] Then $p = \deref{p'}$. \\
%      By inversion, $\tc{\gG}{\lv{x}{p'}}{\own{\gt}}$. \\
%      There are two possible rules for deriving $\tc{\gG}{\lv{x}{\deref{p'}}}{\gt'}$.
%      \begin{itemize}
%	\item[\textsc{PT-DEOWN}]
%	  By inversion, $\tc{\gG}{\lv{x}{p'}}{\own{\gt'}}$. \\
%	  By induction $\own{\gt} =\ \own{\gt'}$. \\
%	  Ergo $\gt = \gt'$.
%	\item[\textsc{PT-DEREF}]
%	  By inversion, $\tc{\gG}{\lv{x}{p'}}{\tyref{\lt}{q}{\gt'}}$. \\
%	  By induction $\own{\gt} = \tyref{\lt}{q}{\gt'}$, a contradiction. \\
%	  Thus, this case is impossible.
%      \end{itemize}
%    \item[\textsc{PT-DEREF}] Then $p = \deref{p'}$. \\
%      Similar to the \textsc{PT-DEOWN} case, 
%      but swapping the roles of \textsc{PT-DEOWN} and \textsc{PT-DEREF}.
%    \item[\textsc{PT-PROJ}] Then $p = \proj{p'}{i}$ and $\gt = \gt_i$. \\
%      By inversion, $\tc{\gG}{\lv{x}{p'}}{\subrec{\gt}{i}{n}}$. \\
%      There is only one rule (\textsc{PT-PROJ}) for deriving $\tc{\gG}{\lv{x}{\proj{p'}{i}}}{\gt'}$. \\
%      Thus, by inversion, $\gt' = \gt'_i$ and $\tc{\gG}{\lv{x}{p'}}{\subrec{\gt'}{i}{n'}}$. \\
%      By induction, $\subrec{\gt}{i}{n} = \subrec{\gt'}{i}{n'}$. \\
%      Ergo $n = n'$ and $\gt_i = \gt'_i$.
%    \item[\textsc{PT-UNROLL}] 
%      Then $p = \unroll{\tyfix{X}{\gt_0}}{p'}$ and $\gt = \sub{X}{\tyfix{X}{\gt_0}}{\gt_0}$. \\
%      There is only one rule (\textsc{PT-UNROLL}) for deriving
%      $\tc{\gG}{\lv{x}{\unroll{\tyfix{X}{\gt_0}}{p'}}}{\gt'}$. \\
%      Thus, by inversion, $\gt' = \sub{X}{\tyfix{X}{\gt_0}}{\gt_0}$. \\
%      Ergo $\gt = \sub{X}{\tyfix{X}{\gt_0}}{\gt_0} = \gt'$.
%  \end{itemize}
%\end{proof}

@@

Due to the extra structure from layouts, addressing memory now requires
more than simple labels. Instead we use a runtime analogue of variables and paths.
Allocations are chunks of memory allocated and freed atomically.
They correspond to either variables or heap allocations.
Routes provide relative specification into layouts similar to paths with two key differences.

@@

The $\base$ route, projection, and unrolling are effectively identical to their path equivalents.
The $\pay{r}$ route refers to the payload $l$ in a variant $\varval{c}{l}$.
This is primarily used for match-by-reference.
Unlike paths, there is no dereference route.
This forces any pointer following into path evaluation (into routes)
rather than doing so while reading memory at a route.
The address placeholder from before is simply a pair of an allocation label and a route.
The exception is that unique pointers needs only an allocation since a unique pointer
always points to a complete heap allocation.
That is:

@@

Our core operations are now defined,
but we still need to describe the typing our runtime data
so that we can properly state and prove our type preservation lemmas.

%With path typing and evaluation now defined, we would like to prove progress
%and preservation for them. Preservation will require we define typing for
%runtime data (routes, cells, and layouts), which involves a slight bit of trickery
%around variant discriminants and the payload route.

%Progress will require a way to guarantee that the necessary \textsc{read}s 
%can be completed, i.e. any pointers we need to dereference are in fact initialized
%(a property we call \emph{shallow initialization}).
%We will accomplish this by using a kind of shadow heap, which will simply track
%initialization rather than values.
%This will require an analogue of \textsc{read} that extracts the initialization
%data for a path, a proof that this operation preserves types 
%(which itselfs requires typing for this shadow heap),
%a coherence check to ensure the shadow heap models the runtime heap,
%and a proof that a shallowly initialized path evaluates to a readable route.

@@

\begin{mathpar}
\mprset{flushleft}
\infer
{ 
\dom{H} = \dom{\gS} \\\\
\forall \ga \in \dom(\gS).~\gS(\ga)~\textrm{closed} \\\\
\forall \ga \in \dom{H}.~\tc{\gS;\gY}{H(\ga)}{\gS(\ga)} \\\\
\forall (\ga,r) \in \dom{\gY}.~\Read{H}{\addr{\ga}{r}}{\varval{\gY(\ga,r)}{l}}
}
{\tc{}{H}{\gS;\gY}}
\end{mathpar}

@@

It can also catch a missing free that would orphan memory.
\begin{verbatim}
{
  let x: ~int = ~3;
  // the shadow of x is `~ int'

  // the stack variable x will be popped off the stack, orphaning the heap memory at *x
  // the type checker can identify this error by seeing that the shadow of x contains `~' data
  // by requiring that shadows cannot contain `~' when freed, the type checker prevents leaks
}
\end{verbatim}

@@

\section*{Shadow Coherence}
Of course, we also need to check that our shadow heap correctly models
the actual heap. For the most part, it is exactly the relation you would expect.
%We require two auxiliary judgments to check for certain initialization states of layouts.
%The $\textsc{init}(H,l)$ judgment checks that $l$ is \emph{fully initialized}, 
%i.e. $\void$ does not occur in any memory reachable from $l$.
%These layouts are safe to use because we cannot access uninitialized memory from them.
The use of \textsc{init} in \textsc{LS-REF} and \textsc{LS-VAR} ensures that
the pointees of borrowed references and the payload of initialized variants
are fully initialized.
The $\textsc{droppable}(l)$ judgment checks that $l$ owns no initialized heap memory,
i.e. $\own{\addr{\ga}{r}}$ is not reachable from $l$ except through references.
These layouts are safe to deallocate because we cannot orphan heap memory by doing so.
The $\tc{V}{H}{\gU}$ judgment ensures that the shadow heap corresponds to the real heap.
\newline

%\fbox{$\textsc{init}(H,l)$}

%\begin{mathpar}
%\infer[I-INT]
%{ }
%{\textsc{init}(H,z)}
%\and
%\infer[I-OWN]
%{\Read{H}{\addr{\ga}{r}}{l} \\ \textsc{init}(H,l)}
%{\textsc{init}(H,\own{\addr{\ga}{r}})}
%\and
%\infer[I-REF]
%{\Read{H}{\addr{\ga}{r}}{l} \\ \textsc{init}(H,l)}
%{\textsc{init}(H,\refval{q}{\addr{\ga}{r}})}
%\and
%\infer[I-VAR]
%{\textsc{init}(H,c) \\ \textsc{init}(H,l)}
%{\textsc{init}(H,\varval{c}{l})}
%\and
%\infer[I-REC]
%{\forall i.~\textsc{init}(H,l_i)}
%{\textsc{init}(H,\subrec{l}{i}{n})}
%\and
%\infer[I-ROLL]
%{\textsc{init}(H,l)}
%{\textsc{init}(H,\roll{\gt}{l})}
%\end{mathpar}

@@

\begin{mathpar}
\mprset{flushleft}
\infer
{ 
\dom{V}=\dom{\gU} \\\\
\forall x \in \dom{V}.~\tc{H}{H(V(x))}{\gU(x)}
}
{\tc{V}{H}{\gU}}
\end{mathpar}

\section*{Shadow Preservation}
The second main preservation lemma for our core operations.
The initialization data, in the form of a shadow,
is preserved by path evaulation and \textsc{read}
so that the resulting layout is correctly described by the shadow.
%We need to tie the shadow of a shallowly initialized path
%to the layout read from a route when the path evaluated to the route.
%This will let us carry guarantees from paths and shadows over to routes and layouts.

>>>>

@@

Due to the extra structure from layouts, addressing memory now requires
more than simple labels. Instead we use a runtime analogue of variables and paths.
Allocations are chunks of memory allocated and freed atomically.
They correspond to either variables or heap allocations.
Routes are like paths except for two key differences.

@@

The $\base$ route, projection, and unrolling are effectively identical to their path equivalents.
The $\pay{r}$ route refers to the payload $l$ in a variant $\varval{c}{l}$.
This is primarily used for match-by-reference.
Unlike paths, there is no dereference route.
This forces any pointer following into path evaluation (into routes)
rather than doing so while reading memory at a route.
The address placeholder from before is simply a pair of an allocation label and a route.
The exception is that unique pointers need only an allocation since a unique pointer
always points to a complete heap allocation.
That is:

@@

Our core operations are now defined,
but we still need to describe the typing of our runtime data
so that we can properly state and prove our type preservation lemmas.

@@

\begin{mathpar}
\mprset{flushleft}
\infer
{\dom{H} = \dom{\gS} \\\\
\forall \ga \in \dom(\gS).~\gS(\ga)~\textrm{closed} \\\\
\forall \ga \in \dom{H}.~\tc{\gS;\gY}{H(\ga)}{\gS(\ga)} \\\\
\forall (\ga,r) \in \dom{\gY}.~\Read{H}{\addr{\ga}{r}}{\varval{\gY(\ga,r)}{l}} \\\\
\forall \ga, \ga', \ga'' \in \dom H.\ \Read{H}{\ga}{\own{\ga''}} \land
				      \Read{H}{\ga'}{\own{\ga''}} \implies \ga = \ga' }
{\tc{}{H}{\gS;\gY}}
\end{mathpar}

@@

It can also catch a missing free that would leak memory.
\begin{verbatim}
{
  let x: ~int = ~3;
  // the shadow of x is `~ int'

  // the stack variable x will be popped off the stack, leaking the heap memory at *x
  // the type checker can identify this error by seeing that the shadow of x contains `~' data
  // by requiring that shadows cannot contain `~' when freed, the type checker prevents leaks
}
\end{verbatim}

@@

\section*{Shadow Coherence}
Of course, we also need to check that our shadow heap correctly models
the actual heap. For the most part, it is exactly the relation you would expect.
The use of \textsc{init} in \textsc{LS-REF} and \textsc{LS-VAR} ensures that
the pointees of borrowed references and the payload of initialized variants
are fully initialized.
The $\textsc{droppable}(l)$ judgment checks that $l$ owns no initialized heap memory,
i.e. $\own{\addr{\ga}{r}}$ is not reachable from $l$ except through references.
These layouts are safe to deallocate because we cannot leak heap memory by doing so.
The $\tc{V}{H}{\gU}$ judgment ensures that the shadow heap corresponds to the real heap.
\newline

@@

\begin{mathpar}
\mprset{flushleft}
\infer
{ 
\dom{V}=\dom{\gU} \\\\
\dom H = \textsc{reachable}(V,H) \\\\
\forall x \in \dom{V}.~\tc{H}{H(V(x))}{\gU(x)}
}
{\tc{V}{H}{\gU}}
\end{mathpar}

\section*{Shadow Preservation}
The second main preservation lemma for our core operations.
The initialization data, in the form of a shadow,
is preserved by path evaulation and \textsc{read}
so that the resulting layout is correctly described by the shadow.

### 2015-02-26 paper/stmts.tex
title, abstract, first round of corrections

\section*{Statements}
We have finally reached the outermost shell of Patina!
The statements layer is primarily about heap manipulation and chaining expressions.
It mostly just relies on judgments we have already defined.
$\nop$ and $\seq{s}{s}$ are simply for chaining together statements.
$\set{\lv{x}{p}}{e}$ is for initializing a path with the result of an expression.
$\new{\lv{x}{p}}{\lv{x}{p}}$ is similar, but allocates the result on the heap 
and initializes the path with a new unique pointer. 
Treating this as a statement rather than an expression
simplifies the expression preservation and progress lemmas.
Note that both versions are for \emph{initialization} and not \emph{assignment}.
We only allow initialization of droppable paths, which we guarantee will not orphan memory.
$\swap{\lv{x}{p}}{\lv{x}{p}}$ will perform a shallow swap of two writable paths.
This is the only way to actually mutate through a mutable borrow since we do not allow
interior of a borrowed reference to be uninitialized even temporarily.
$\free{\lv{x}{p}}$ performs a shallow free and is the inverse of \textsf{new}.
$\push{\lt}{x}{\gt}{s}$ will create a new stack variable.
$\pop{x}{s}$ will pop that stack variable once the statement that requires it finishes.
It is for evaluation purposes and should not appear in user code.

@@

Initializing with a value requires that the path be droppable so that nothing will be orphaned.
It also requires the expression type to be a subtype of the path type.
As discussed above, this refers to lifetimes: the subtype lifetime is greater than or
equal to the lifetime of the supertype. This requirement guarantees that the value
stored at the path will be valid for at least as long as the path is valid.
Finally, we initialize the path in the output shadow heap.

@@

Pushing a new stack variable involves checking that the body is well typed
under the appropriately extended contexts. Newly pushed variables start uninitialized.
By the end of the body, the variable must be droppable so that popping it will not
orphan memory. When checking the body, we extend the lifetime relation with the new lifetime
by adding the new lifetime as the smallest lifetime in the relation (it is related to everything
else). Afterwards, we must filter everything involving this lifetime from the shadow heap.
This should only consist of removing loans from banks. We also "deallocate" the stack variable
from the shadow heap by restricting the domain to the domain of the input.

@@

\subsection*{Statement Progress}
Statement progress is essentially the assertion that a well-typed statement will
never attempt to read uninitialized memory, free borrowed memory, and so on.
Essentially, that well-typed programs have sufficient memory safety to take a step.
The restriction to statements that end with an empty shadow heap allows us
to avoid the cases with orphan memory.

>>>>

\section*{Statements}
We have finally reached the outermost shell of Patina!
The statements layer is primarily about heap manipulation and chaining expressions.
It mostly just relies on judgments we have already defined.
$\nop$ and $\seq{s}{s}$ are simply for chaining together statements.
$\set{\lv{x}{p}}{e}$ is for initializing a path with the result of an expression.
$\new{\lv{x}{p}}{\lv{x}{p}}$ is similar, but allocates the result on the heap 
and initializes the path with a new unique pointer. 
Treating this as a statement rather than an expression
simplifies the expression preservation and progress lemmas.
Note that both versions are for \emph{initialization} and not \emph{assignment}.
We only allow initialization of droppable paths, which we guarantee will not leak memory.
$\swap{\lv{x}{p}}{\lv{x}{p}}$ will perform a shallow swap of two writable paths.
This is the only way to actually mutate through a mutable borrow since we do not allow
interior of a borrowed reference to be uninitialized even temporarily.
$\free{\lv{x}{p}}$ performs a shallow free and is the inverse of \textsf{new}.
$\push{\lt}{x}{\gt}{s}$ will create a new stack variable.
$\pop{x}{s}$ will pop that stack variable once the statement that requires it finishes.
It is for evaluation purposes and should not appear in user code.

@@

Initializing with a value requires that the path be droppable so that nothing will be leaked.
It also requires the expression type to be a subtype of the path type.
As discussed above, this refers to lifetimes: the subtype lifetime is greater than or
equal to the lifetime of the supertype. This requirement guarantees that the value
stored at the path will be valid for at least as long as the path is valid.
Finally, we initialize the path in the output shadow heap.

@@

Pushing a new stack variable involves checking that the body is well typed
under the appropriately extended contexts. Newly pushed variables start uninitialized.
By the end of the body, the variable must be droppable so that popping it will not
leak memory. When checking the body, we extend the lifetime relation with the new lifetime
by adding the new lifetime as the smallest lifetime in the relation (it is related to everything
else). Afterwards, we must filter everything involving this lifetime from the shadow heap.
This should only consist of removing loans from banks. We also "deallocate" the stack variable
from the shadow heap by restricting the domain to the domain of the input.

@@

\subsection*{Statement Progress}
Statement progress is essentially the assertion that a well-typed statement will
never attempt to read uninitialized memory, free borrowed memory, and so on.
Essentially, that well-typed programs have sufficient memory safety to take a step.
The restriction to statements that end with an empty shadow heap allows us
to avoid the cases with leaked memory.
